\section{Python}

\idx{Python}Script\idx{Script Language} language which wont bite too much. It won't do concurrency\footnote{Unless you \textsl{\textbf{really}} know what you are doing, and 95\% of those who think they do don't.}, syntax errors are produced at runtime, and it is numerically pretty slow. If you can accept this (e.g., if you are doing script work) it's a pretty good language.

\subsection{Versions}

\subsection{Embedded Systems}

\idxx{Python!Embedded systems}{Embedded systems}As a stop-the-world\idxx{Garbage collection!Stop-the-world}{Stop-the-world Garbage collection} garbage collected language, Python is inherently problematic for embedded applications. To ensure successful communication with peripheral devices (e.g., radios, flash chips and digital sensors) that employ time-critical communication protocols\idxx{Communication!Time critical}{Time-critical communicaiton}, one is likely to have to either disable garbage collection during such operations or force it just before. This will pollute the codebase, introduce potential and hard-to-find bugs, and result in an efficiency penalty.

MicroPython\idx{MicroPython}\footnote{\url{https://micropython.org}} is an implementation of Python 3 with microcontrollers in mind. It is efficient for quick and dirty prototyping, but shouldn't be used for production purposes.

\subsection{Embedding}

% purpose: access to python modules, plugin/scripting language

% c api
\idxx{Python!Embedding}{Embedding}\url{https://docs.python.org/3/c-api/intro.html}

% binding to other languages (e.g., go)

\subsection{Symbolic Mathematics}

\idx{SymPy}\url{https://www.sympy.org}

Example script:
\begin{minted}[breaklines]{python}
import sympy

s = input('Please enter an expression over i and j: ')
expr = sympy.sympify(s)
symi = sympy.symbols('i')
symj = sympy.symbols('j')

print('expr='+str(expr))
print('diff(expr, i)='+str(sympy.diff(expr, symi)))

for i in range(10):
    j = i % 3
    v = expr.subs([(symi, i), (symj, j)])
    print('i=%d j=%d => result=%f' % (i, j, v))
\end{minted}

Example session of running script:
\begin{verbatim}
Please enter an expression over i and j: 3*i*i+4.83*i+j 
expr=3*i**2 + 4.83*i + j
diff(expr, i)=6*i + 4.83
i=0 j=0 => result=0.000000
i=1 j=1 => result=8.830000
i=2 j=2 => result=23.660000
i=3 j=0 => result=41.490000
i=4 j=1 => result=68.320000
i=5 j=2 => result=101.150000
i=6 j=0 => result=136.980000
i=7 j=1 => result=181.810000
i=8 j=2 => result=232.640000
i=9 j=0 => result=286.470000
\end{verbatim}

\subsection{Object Orientation}

\subsection{AsyncIO}

